A unified pseudo-$C_\ell$ framework

### 2018-09-04 doc/Paper/main.tex
started implementing second round of comments

    While maximum-likelihood~\citep{1998PhRvD..57.2117B,2003PhRvD..67b3001W} or minimum-variance and quadratic estimators~\citep{1997PhRvD..55.5895T,2000MNRAS.317L..23H,2001PhRvD..64f3001T} can recover the power spectrum with virtually no loss of information (see also the approaches of e.g. \cite{2004ApJS..155..227E,2008MNRAS.389.1284T,2016MNRAS.455.4452A}, based on Bayesian sampling methods), their computational implementation is prohibitively expensive for high resolution data, since these methods scale as $\ell_{\rm max}^6$ with the maximum multipole $\ell_{\rm max}$ (or $\ell_{\rm max}^4$ for optimal algorithms such as that of \cite{2003PhRvD..67b3001W}). For this reason, pseudo-$C_\ell$ algorithms \citep{1973ApJ...185..413P,2001PhRvD..64h3003W,2001ApJ...561L..11S,2001PhRvD..64h3003W,2002ApJ...567....2H,2002MNRAS.336.1304H,2004MNRAS.350..914C} have become a popular alternative that preserves the $\ell_{\rm max}^3$ scaling of spherical harmonic transforms (and faster in the flat-sky approximation). This method can be used on fields with arbitrary spin \citep{2003ApJS..148..161K} in both curved and flat skies, has been extended to deal with contaminant deprojection in scalar fields \citep{2017MNRAS.465.1847E}, and can be optimized to measure the CMB $E/B$-mode power spectrum \citep{2002PhRvD..65b3505L,2003PhRvD..67b3501B,2006PhRvD..74h3002S,2009PhRvD..79l3515G,2011PhRvD..83h3003B}. Unfortunately, in spite of the advantages of this method, while a number of power spectrum codes solving some of these problems have been put forward, no public, validated and well-maintained code exists supporting all of these capabilities consistently. This paper presents a public software package, \nmt, that unifies and extends all of these features providing a general framework to estimate pseudo-$C_\ell$ angular power spectra in a wide variety of scenarios.

@@

      Finally we define the matrix operator $\hat{\sf Y}^s_{\ell m}$ to have $_s{\bf Y}^{E,B}_{\ell m}$ as columns:
      \begin{equation}
        \hat{\sf Y}^s_{\ell m}\equiv\left(_s{\bf Y}^E_{\ell m},_s{\bf Y}^B_{\ell m},\right).
      \end{equation}
      These matrices satisfy the following properties:
      \begin{align}
        &\hat{\sf Y}^{s\dag}_{\ell m}=(-1)^{m+s}\hat{\sf Y}^{-s}_{\ell -m}\\
        &\int d\nv \hat{\sf Y}^s_{\ell m}\hat{\sf Y}^{s\dag}_{\ell'm'}=\hat\mI\delta_{\ell\ell'}\delta_{mm'},\\
        &\int d\nv \left(\hat{\sf Y}^{s\dag}_{\bf l}(\nv)\hat{\sf Y}^{s}_{{\bf l}_1}(\nv)\right)\,\hat{\sf Y}^0_{{\bf l}_2}(\nv)\equiv\hat{\sf D}^s_{{\bf l}{\bf l}_1{\bf l}_2},
      \end{align}
      where we have abbreviated the pair $(\ell,m)$ as ${\bf l}$, and
      \begin{align}\label{eq:dmat}
        \hat{\sf D}^s_{{\bf l}{\bf l}_1{\bf l}_2}&=(-1)^{s+m}\sqrt{\frac{(2\ell+1)(2\ell_1+1)(2\ell_2+1)}{4\pi}}\\\nonumber
        &\hspace{10pt}\wtj{\ell}{\ell_1}{\ell_2}{-m}{m_1}{m_2}
        \wtj{\ell}{\ell_1}{\ell_2}{s}{-s}{0}\,\hat{\sf d}_{\ell+\ell_1+\ell_2}.
      \end{align}
      Here
      \begin{equation}
        \hat{\sf d}_n= \frac{1}{2}\left(\begin{array}{cc}
                                            1+(-1)^n & -i[1-(-1)^n]\\
                                            i[1-(-1)^n] & 1+(-1)^n
                                          \end{array}\right),
      \end{equation}
      and the Wigner 3$j$ symbols satisfy the orthogonality relation
      \begin{equation}\label{eq:3jorth}
        \sum_{mm_1}\wtj{\ell}{\ell_1}{\ell_2}{m}{m_1}{m_2}\wtj{\ell}{\ell_1}{\ell_3}{m}{m_1}{m_3}=\frac{\delta_{\ell_2\ell_3}\delta_{m_2m_3}}{2\ell_2+1}
      \end{equation}

>>>>

    While maximum-likelihood~\citep{1998PhRvD..57.2117B,2003PhRvD..67b3001W} or minimum-variance and quadratic estimators~\citep{1997PhRvD..55.5895T,2000MNRAS.317L..23H,2001PhRvD..64f3001T} can recover the power spectrum with virtually no loss of information (see also the approaches of e.g. \cite{2004ApJS..155..227E,2008MNRAS.389.1284T,2016MNRAS.455.4452A}, based on Bayesian sampling methods), their computational implementation is prohibitively expensive for high resolution data, since these methods scale as $\ell_{\rm max}^6$ with the maximum multipole $\ell_{\rm max}$ (or $\ell_{\rm max}^4$ for optimal algorithms such as that of \cite{2003PhRvD..67b3001W}). For this reason, pseudo-$C_\ell$ algorithms \citep{1973ApJ...185..413P,2001PhRvD..64h3003W,2001ApJ...561L..11S,2001PhRvD..64h3003W,2002ApJ...567....2H,2002MNRAS.336.1304H,2004MNRAS.350..914C} have become a popular alternative that preserves the $\ell_{\rm max}^3$ scaling of spherical harmonic transforms (and faster in the flat-sky approximation). This method can be used on fields with arbitrary spin \citep{2003ApJS..148..161K} in both curved and flat skies, has been extended to deal with contaminant deprojection in scalar fields \citep{2017MNRAS.465.1847E}, and can be optimized to measure the CMB $E/B$-mode power spectrum \citep{2002PhRvD..65b3505L,2003PhRvD..67b3501B,2006PhRvD..74h3002S,2009PhRvD..79l3515G,2011PhRvD..83h3003B}. Unfortunately, in spite of the advantages of this method, while a number of power spectrum codes solving some of these problems have been put forward, no public, validated and well-maintained code exists supporting all of these capabilities consistently. This paper presents a public software package, \nmt, that unifies all of these features providing a general framework to estimate pseudo-$C_\ell$ angular power spectra in a wide variety of scenarios. The paper also presents a few novel results that were derived as part of the development of \nmt. These include extending the contaminant deprojection formalism to spin-$2$ fields, combining it with $E/B$ purification and providing a consistent derivation of the same functionality in the flat-sky approximation.

@@

      Finally we define the matrix operator $\hat{\sf Y}^s_{\ell m}$ to have $_s{\bf Y}^{E,B}_{\ell m}$ as columns:
      \begin{equation}
        \hat{\sf Y}^s_{\ell m}\equiv\left(_s{\bf Y}^E_{\ell m},_s{\bf Y}^B_{\ell m},\right).
      \end{equation}
      These matrices satisfy the following properties:
      \begin{align}
        &\hat{\sf Y}^{s\dag}_{\ell m}=(-1)^{m+s}\hat{\sf Y}^{-s}_{\ell -m}\\
        &\int d\nv\,\hat{\sf Y}^{s\dag}_{\ell m}\hat{\sf Y}^s_{\ell'm'}=\hat\mI\delta_{\ell\ell'}\delta_{mm'},\\
        &\int d\nv \left(\hat{\sf Y}^{s\dag}_{\bf l}(\nv)\hat{\sf Y}^{s}_{{\bf l}_1}(\nv)\right)\,\hat{\sf Y}^0_{{\bf l}_2}(\nv)\equiv\hat{\sf D}^s_{{\bf l}{\bf l}_1{\bf l}_2},
      \end{align}
      where we have abbreviated the pair $(\ell,m)$ as ${\bf l}$, and
      \begin{align}\label{eq:dmat}
        \hat{\sf D}^s_{{\bf l}{\bf l}_1{\bf l}_2}&=(-1)^{s+m}\sqrt{\frac{(2\ell+1)(2\ell_1+1)(2\ell_2+1)}{4\pi}}\\\nonumber
        &\hspace{10pt}\wtj{\ell}{\ell_1}{\ell_2}{-m}{m_1}{m_2}
        \wtj{\ell}{\ell_1}{\ell_2}{s}{-s}{0}\,\hat{\sf d}_{\ell+\ell_1+\ell_2}.
      \end{align}
      Here
      \begin{equation}
        \hat{\sf d}_n= \frac{1}{2}\left(\begin{array}{cc}
                                            1+(-1)^n & -i[1-(-1)^n]\\
                                            i[1-(-1)^n] & 1+(-1)^n
                                          \end{array}\right),
      \end{equation}
      and the Wigner 3$j$ symbols satisfy the orthogonality relation
      \begin{equation}\label{eq:3jorth}
        \sum_{mm_1}\wtj{\ell}{\ell_1}{\ell_2}{m}{m_1}{m_2}\wtj{\ell}{\ell_1}{\ell_3}{m}{m_1}{m_3}=\frac{\delta_{\ell_2\ell_3}\delta_{m_2m_3}}{2\ell_2+1}
      \end{equation}
